---
title: "Bayesian Final Project"
output: html_document
date: "2023-11-07"
---

```{r}


setwd("~/Desktop")

Swedish <- read.csv("/Users/matthewludwig/Desktop/SwedishMotorInsurance.csv", header=TRUE, as.is=TRUE, sep=',')

head(Swedish) 
summary(Swedish)
Swedish$Zone <- as.factor(Swedish$Zone)


class(Swedish)
```

# Split Data into 7 based on Geographic Zone:

```{r}

Zone1 <- Swedish[Swedish$Zone==1, ]
Zone2 <- Swedish[Swedish$Zone==2, ]
Zone3 <- Swedish[Swedish$Zone==3, ]
Zone4 <- Swedish[Swedish$Zone==4, ]
Zone5 <- Swedish[Swedish$Zone==5, ]
Zone6 <- Swedish[Swedish$Zone==6, ]
Zone7 <- Swedish[Swedish$Zone==7, ]
```

# Summary Statistics:

# Zone 1:

```{r}

par(mfrow = c(2, 3))  


boxplot(Claims/Insured ~ Kilometres, data=Zone1, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone1, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone1, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone1, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone1, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone1, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone1)
```

# Zone 2:

```{r}

par(mfrow = c(2, 3))  


boxplot(Claims/Insured ~ Kilometres, data=Zone2, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone2, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone2, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone2, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone2, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone2, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone2)

```

# Zone 3:

```{r}

par(mfrow = c(2, 3))  


boxplot(Claims/Insured ~ Kilometres, data=Zone3, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone3, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone3, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone3, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone3, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone3, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone3)
```

# Zone 4:

```{r}

par(mfrow = c(2, 3))  # 2 rows, 3 columns

boxplot(Claims/Insured ~ Kilometres, data=Zone4, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone4, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone4, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone4, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone4, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone4, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone4)
```

# Zone 5:

```{r}

par(mfrow = c(2, 3))  # 2 rows, 3 columns

boxplot(Claims/Insured ~ Kilometres, data=Zone5, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone5, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone5, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone5, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone5, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone5, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone5)
```

# Zone 6:

```{r}

par(mfrow = c(2, 3))  # 2 rows, 3 columns

boxplot(Claims/Insured ~ Kilometres, data=Zone6, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone6, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone6, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone6, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone6, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone6, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone6)
```

# Zone 7:

```{r}

par(mfrow = c(2, 3))  # 2 rows, 3 columns

boxplot(Claims/Insured ~ Kilometres, data=Zone7, xlab="Distance Driven",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Bonus, data=Zone7, xlab="Accident Free Years",
        ylab="Average Claim Number", ylim = c(0, 0.6))

boxplot(Claims/Insured ~ Make, data=Zone7, xlab="Auto Make",
        ylab="Average Claim Number", ylim = c(0, 0.6))


boxplot(Payment ~ Kilometres, data=Zone7, xlab="Distance Driven",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Bonus, data=Zone7, xlab="Accident Free Years",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

boxplot(Payment ~ Make, data=Zone7, xlab="Auto Make",
        ylab="Average Claim Severity", ylim = c(0, 1000000))

summary(Zone7)
```

# Histograms of Claim Frequencies in Zones

```{r}

options(scipen = 999)
par(mfrow=c(2, 2)) 




hist(Zone1$Claims, main = "Claim Frequenices in Zone 1", xlab = "Number of Claims", col = "lightcoral", border = "black", breaks = 500, xlim = c(0,100))


hist(Zone2$Claims, main = "Claim Frequencies in Zone 2", xlab = "Number of Claims", col = "lightcoral", border = "black", breaks = 500, xlim = c(0,100))

hist(Zone3$Claims, breaks = 500, main = "Claim Frequencies in Zone 3", xlab = "Number of Claims", col = "skyblue", border = "black", xlim = c(0,150))

hist(Zone4$Claims, main = "Claim Frequencies in Zone 4", xlab = "Number of Claims", col = "lightgreen", border = "black", breaks = 500, xlim = c(0,100))

hist(Zone5$Claims, main = "Claim Frequencies in Zone 5", xlab = "Number of Claims", col = "lightpink", border = "black", breaks = 500, xlim = c(0,100))

hist(Zone6$Claims, main = "Claim Frequencies in Zone 6", xlab = "Number of Claims", col = "lightcoral", border = "black", breaks = 500, xlim = c(0,100))

hist(Zone7$Claims, main = "Claim Frequencies in Zone 7", xlab = "Number of Claims", col = "lightcoral", border = "black", breaks = 500, xlim = c(0,100))


```

# Histograms of Claim Severity in Zones

```{r}
options(scipen = 999)
par(mfrow=c(2, 2)) 




hist(Zone1$Payment, main = "Zone 1 Severity", xlab = "Severity", col = "lightcoral", border = "black", breaks = 200, xlim = c(0,1000000))


hist(Zone2$Payment, main = "Zone 2 Severity", xlab = "Severity", col = "lightcoral", border = "black", breaks = 200, xlim = c(0,1000000))

hist(Zone3$Payment, breaks = 200, main = "Zone 3 Severity", xlab = "Severity", col = "skyblue", border = "black", xlim = c(0,1000000))

hist(Zone4$Payment, main = "Zone 4 Severity", xlab = "Severity", col = "lightgreen", border = "black", breaks = 200, xlim = c(0,1000000))

hist(Zone5$Payment, main = "Zone 5 Severity", xlab = "Severity", col = "lightpink", border = "black", breaks = 200, xlim = c(0,1000000))

hist(Zone6$Payment, main = "Zone 6 Severity", xlab = "Severity", col = "lightcoral", border = "black", breaks = 200, xlim = c(0,1000000))

hist(Zone7$Payment, main = "Zone 7 Severity", xlab = "Severity", col = "lightcoral", border = "black", breaks = 200, xlim = c(0,1000000))




```

\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~

# Claim Frequency Poisson Hierarchical Model:

```{r}


#Poisson-gamma hierarchical model
#rate parameter for Poisson distribution of claim frequency drawn from gamma distribution
#Gamma distribution


#Stan code used to fit and find parameters for the hierarchical model:

library(rstan)

stan_model_code <- '

data {
  int<lower=0> N1; //# observations in Zone 1
  int<lower=0> N2; 
  int<lower=0> N3; 
  int<lower=0> N4; 
  int<lower=0> N5; 
  int<lower=0> N6; 
  int<lower=0> N7; 
  
  int<lower=0> y1[N1]; //frequencies in Zone 1
  int<lower=0> y2[N2]; 
  int<lower=0> y3[N3]; 
  int<lower=0> y4[N4]; 
  int<lower=0> y5[N5]; 
  int<lower=0> y6[N6]; 
  int<lower=0> y7[N7]; 
}



parameters {
  real<lower=0> alpha; //shape for gamma distribution
  real<lower=0> beta;  //rate for gamma distribution
  
  real<lower=0> lambda1; //poisson rate parameter for Zone 1
  real<lower=0> lambda2; // 
  real<lower=0> lambda3; 
  real<lower=0> lambda4; 
  real<lower=0> lambda5; 
  real<lower=0> lambda6; 
  real<lower=0> lambda7; 
}




model {

  //priors
  alpha ~ gamma(1, .01); // prior for alpha
  beta ~ gamma(1,.01);  // prior for beta
  
  
  
  //hierarchical model
  lambda1 ~ gamma(alpha, beta);
  lambda2 ~ gamma(alpha, beta);
  lambda3 ~ gamma(alpha, beta);
  lambda4 ~ gamma(alpha, beta);
  lambda5 ~ gamma(alpha, beta);
  lambda6 ~ gamma(alpha, beta);
  lambda7 ~ gamma(alpha, beta);
  
  
  
  //likelihoods
  y1 ~ poisson(lambda1);
  y2 ~ poisson(lambda2);
  y3 ~ poisson(lambda3);
  y4 ~ poisson(lambda4);
  y5 ~ poisson(lambda5);
  y6 ~ poisson(lambda6);
  y7 ~ poisson(lambda7);
}


generated quantities {
  int<lower=0> y_rep1[N1];
  int<lower=0> y_rep2[N2];
  int<lower=0> y_rep3[N3];
  int<lower=0> y_rep4[N4];
  int<lower=0> y_rep5[N5];
  int<lower=0> y_rep6[N6];
  int<lower=0> y_rep7[N7];
  
  for (i in 1:N1) y_rep1[i] = poisson_rng(lambda1);
  for (i in 1:N2) y_rep2[i] = poisson_rng(lambda2);
  for (i in 1:N3) y_rep3[i] = poisson_rng(lambda3);
  for (i in 1:N4) y_rep4[i] = poisson_rng(lambda4);
  for (i in 1:N5) y_rep5[i] = poisson_rng(lambda5);
  for (i in 1:N6) y_rep6[i] = poisson_rng(lambda6);
  for (i in 1:N7) y_rep7[i] = poisson_rng(lambda7);
}

'

Frequencies1 <- Zone1$Claims
Frequencies2 <- Zone2$Claims
Frequencies3 <- Zone3$Claims
Frequencies4 <- Zone4$Claims
Frequencies5 <- Zone5$Claims
Frequencies6 <- Zone6$Claims
Frequencies7 <- Zone7$Claims



# Compile the Stan model
stan_model <- stan_model(model_code = stan_model_code)

data_list <- list(
  N1 = length(Frequencies1),
  y1 = Frequencies1,
  N2 = length(Frequencies2),
  y2 = Frequencies2,
  N3 = length(Frequencies3),
  y3 = Frequencies3,
  N4 = length(Frequencies4),
  y4 = Frequencies4,
  N5 = length(Frequencies5),
  y5 = Frequencies5,
  N6 = length(Frequencies6),
  y6 = Frequencies6,
  N7 = length(Frequencies7),
  y7 = Frequencies7
)

stan_fit <- sampling(stan_model, data = data_list, chains = 4, iter = 2000)
print(stan_fit)

```

Fit model and evaluate:

```{r}

library(rstan)

plot(stan_fit)
traceplot(stan_fit)
summary(stan_fit)


library(bayesplot)
autocorrelation_plot <- mcmc_acf(stan_fit)
#autocorrelation_plot
```

# Posterior Predictive Check:

```{r}

library(bayesplot)

# Extract simulated data from the stan_fit object
yrep1 <- as.array(stan_fit, "y_rep1")
yrep2 <- as.array(stan_fit, "y_rep2")
yrep3 <- as.array(stan_fit, "y_rep3")
yrep4 <- as.array(stan_fit, "y_rep4")
yrep5 <- as.array(stan_fit, "y_rep5")
yrep6 <- as.array(stan_fit, "y_rep6")
yrep7 <- as.array(stan_fit, "y_rep7")

Frequencies1
Frequencies2
Frequencies3
Frequencies4
Frequencies5
Frequencies6
Frequencies7

yrep1fixed <- yrep1[1:6, 1,]
ppc_hist(Frequencies1, yrep1fixed, binwidth = 50)
ppc_dens_overlay(Frequencies1, yrep1fixed, binwidth = 50)

yrep1fixed <- yrep1[1:500, 1,]
ppc_stat(Frequencies1,yrep1fixed, stat='min')
ppc_stat(Frequencies1,yrep1fixed, stat='max')
ppc_stat(Frequencies1,yrep1fixed, stat='min')
ppc_stat(Frequencies1,yrep1fixed)










par(mfrow = c(3, 2))  # 3 rows, 2 columns

# Plot for yrep2
yrep2fixed <- yrep2[1:6, 1,]
ppc_hist(Frequencies2, yrep2fixed, binwidth = 50)
ppc_dens_overlay(Frequencies2, yrep2fixed, binwidth = 50)

# Plot for yrep3
yrep3fixed <- yrep3[1:6, 1,]
ppc_hist(Frequencies3, yrep3fixed, binwidth = 50)
ppc_dens_overlay(Frequencies3, yrep3fixed, binwidth = 50)

# Plot for yrep4
yrep4fixed <- yrep4[1:6, 1,]
ppc_hist(Frequencies4, yrep4fixed, binwidth = 50)
ppc_dens_overlay(Frequencies4, yrep4fixed, binwidth = 50)

# Plot for yrep5
yrep5fixed <- yrep5[1:6, 1,]
ppc_hist(Frequencies5, yrep5fixed, binwidth = 50)
ppc_dens_overlay(Frequencies5, yrep5fixed, binwidth = 50)

# Plot for yrep6
yrep6fixed <- yrep6[1:6, 1,]
ppc_hist(Frequencies6, yrep6fixed, binwidth = 50)
ppc_dens_overlay(Frequencies6, yrep6fixed, binwidth = 50)

# Plot for yrep7
yrep7fixed <- yrep7[1:6, 1,]
ppc_hist(Frequencies7, yrep7fixed, binwidth = 5)
ppc_dens_overlay(Frequencies7, yrep7fixed, binwidth = 5)

# Reset the layout to a single plot
par(mfrow = c(1, 1))


```

# Calculate MSE:

```{r}

# Define a function to calculate MSE
calculate_mse <- function(observed, replicated) {
  mean((replicated - observed)^2, na.rm = TRUE)
}

# Calculate MSE for each zone
mse_zone1 <- calculate_mse(Frequencies1, yrep1fixed)
mse_zone2 <- calculate_mse(Frequencies2, yrep2fixed)
mse_zone3 <- calculate_mse(Frequencies3, yrep3fixed)
mse_zone4 <- calculate_mse(Frequencies4, yrep4fixed)
mse_zone5 <- calculate_mse(Frequencies5, yrep5fixed)
mse_zone6 <- calculate_mse(Frequencies6, yrep6fixed)
mse_zone7 <- calculate_mse(Frequencies7, yrep7fixed)

# Print MSE for each zone
cat("MSE Zone 1:", mse_zone1, "\n")
cat("MSE Zone 2:", mse_zone2, "\n")
cat("MSE Zone 3:", mse_zone3, "\n")
cat("MSE Zone 4:", mse_zone4, "\n")
cat("MSE Zone 5:", mse_zone5, "\n")
cat("MSE Zone 6:", mse_zone6, "\n")
cat("MSE Zone 7:", mse_zone7, "\n")




# Create a bar plot with zone labels
barplot(mse_values, names.arg = zone_labels, col = "skyblue", main = "MSE for Each Zone",
        ylab = "Mean Squared Error (MSE)", ylim = range(0, mse_values * 1.2))



```

# Model the Frequency of Claims with a Hierarchical Poisson Regression Model:

```{r}

#install.packages("rstan")
library(rstan)




stan_code_2 <- '

data {

  int<lower=1> K;                 // Number of predictors

  int<lower=0> N1; //# observations in Zone 1
  int<lower=0> N2; 
  int<lower=0> N3; 
  int<lower=0> N4; 
  int<lower=0> N5; 
  int<lower=0> N6; 
  int<lower=0> N7; 
  
  int<lower=0> y1[N1]; //frequencies in Zone 1
  int<lower=0> y2[N2]; 
  int<lower=0> y3[N3]; 
  int<lower=0> y4[N4]; 
  int<lower=0> y5[N5]; 
  int<lower=0> y6[N6]; 
  int<lower=0> y7[N7]; 
  
  
  int<lower=0> Kilometres1[N1]; //Kilometers in Zone 1
  int<lower=0> Kilometres2[N2];
  int<lower=0> Kilometres3[N3];
  int<lower=0> Kilometres4[N4];
  int<lower=0> Kilometres5[N5];
  int<lower=0> Kilometres6[N6];
  int<lower=0> Kilometres7[N7];
  
  int<lower=1, upper=7> Zones1[N1];   // Geographical zones
  int<lower=1, upper=7> Zones2[N2];
  int<lower=1, upper=7> Zones3[N3];
  int<lower=1, upper=7> Zones4[N4];
  int<lower=1, upper=7> Zones5[N5];
  int<lower=1, upper=7> Zones6[N6];
  int<lower=1, upper=7> Zones7[N7];
  
  int<lower=1> Bonus1[N1]; // No claim bonus
  int<lower=1> Bonus2[N2];
  int<lower=1> Bonus3[N3]; 
  int<lower=1> Bonus4[N4]; 
  int<lower=1> Bonus5[N5]; 
  int<lower=1> Bonus6[N6]; 
  int<lower=1> Bonus7[N7]; 
  
  int<lower=1, upper=9> Make1[N1];   // Make of the vehicle
  int<lower=1, upper=9> Make2[N2];
  int<lower=1, upper=9> Make3[N3];
  int<lower=1, upper=9> Make4[N4];
  int<lower=1, upper=9> Make5[N5];
  int<lower=1, upper=9> Make6[N6];
  int<lower=1, upper=9> Make7[N7];
  
  real<lower=0> Insured1[N1];        // Insured
  real<lower=0> Insured2[N2];
  real<lower=0> Insured3[N3];
  real<lower=0> Insured4[N4];
  real<lower=0> Insured5[N5];
  real<lower=0> Insured6[N6];
  real<lower=0> Insured7[N7];
  
}


parameters {

  real alpha1;               // Intercept
  real beta1[K];             // Coefficients for predictors
  real<lower=0> gamma1;   // Random effects for Zone
  
  real alpha2;               // Intercept
  real beta2[K];             // Coefficients for predictors
  real<lower=0> gamma2;   // Random effects for Zone

  real alpha3;               // Intercept
  real beta3[K];             // Coefficients for predictors
  real<lower=0> gamma3;   // Random effects for Zone
  
  real alpha4;               // Intercept
  real beta4[K];             // Coefficients for predictors
  real<lower=0> gamma4;   // Random effects for Zone
  
  real alpha5;               // Intercept
  real beta5[K];             // Coefficients for predictors
  real<lower=0> gamma5;   // Random effects for Zone
  
  real alpha6;               // Intercept
  real beta6[K];             // Coefficients for predictors
  real<lower=0> gamma6;   // Random effects for Zone
  
  real alpha7;               // Intercept
  real beta7[K];             // Coefficients for predictors
  real<lower=0> gamma7;   // Random effects for Zone


  //real<lower=0> lambda1; //poisson rate parameter for Zone 1
  //real<lower=0> lambda2; // 
  //real<lower=0> lambda3; 
  //real<lower=0> lambda4; 
  //real<lower=0> lambda5; 
  //real<lower=0> lambda6; 
  //real<lower=0> lambda7; 
  
}




// beta[1] = Kilometres
// beta[2] = Zone
// beta[3] = Bonus
// beta[4] = Make
// beta[5] = Insured




model {


// priors

  alpha1 ~ uniform(.001, 0.1);
  beta1 ~ uniform(.001, 1);
  gamma1 ~ uniform(.001, 0.1);

  alpha2 ~ uniform(.001, 0.1);
  beta2 ~ uniform(0.001, 1);
  gamma2 ~ uniform(0.001, 1);

  alpha3 ~ uniform(0.001, 0.1);
  beta3 ~ uniform(0.001, 1);
  gamma3 ~ uniform(0.001, 1);

  alpha4 ~ uniform(0.001, 0.1);
  beta4 ~ uniform(0.001, 0.01);
  gamma4 ~ uniform(0.001, 1);

  alpha5 ~ uniform(0.001, 0.1);
  beta5 ~ uniform(0.001, 1);
  gamma5 ~ uniform(0.00001, 0.0001);
  
  
  
  
  // (lambdas should be between 0-100)
  // linear predictors should be between 0 - 5
  

  // Likelihood for the main model
for (i in 1:N1) {
  real lambda1 = exp(alpha1 + beta1[1] * Kilometres1[i] + beta1[2] * Zones1[i] + beta1[3] * Bonus1[i] + beta1[4] * Make1[i] + beta1[5] * Insured1[i] + gamma1);
  y1[i] ~ poisson(lambda1);
}

for (i in 1:N2) {
  real lambda2 = exp(alpha2 + beta2[1] * Kilometres2[i] + beta2[2] * Zones2[i] +
                    beta2[3] * Bonus2[i] + beta2[4] * Make2[i] + beta2[5] * Insured2[i] +
                    gamma2);
  y2[i] ~ poisson(lambda2);
}

for (i in 1:N3) {
  real lambda3 = exp(alpha3 + beta3[1] * Kilometres3[i] + beta3[2] * Zones3[i] +
                    beta3[3] * Bonus3[i] + beta3[4] * Make3[i] + beta3[5] * Insured3[i] +
                    gamma3);
  y3[i] ~ poisson(lambda3);
}

for (i in 1:N4) {
  real lambda4 = exp(alpha4 + beta4[1] * Kilometres4[i] + beta4[2] * Zones4[i] +
                    beta4[3] * Bonus4[i] + beta4[4] * Make4[i] + beta4[5] * Insured4[i] +
                    gamma4);
  y4[i] ~ poisson(lambda4);
}

for (i in 1:N5) {
  real lambda5 = exp(alpha5 + beta5[1] * Kilometres5[i] + beta5[2] * Zones5[i] +
                    beta5[3] * Bonus5[i] + beta5[4] * Make5[i] + beta5[5] * Insured5[i] +
                    gamma5);
  y5[i] ~ poisson(lambda5);
}

for (i in 1:N6) {
  real lambda6 = exp(alpha6 + beta6[1] * Kilometres6[i] + beta6[2] * Zones6[i] +
                    beta6[3] * Bonus6[i] + beta6[4] * Make6[i] + beta6[5] * Insured6[i] +
                    gamma6);
  y6[i] ~ poisson(lambda6);
}

for (i in 1:N7) {
  real lambda7 = exp(alpha7 + beta7[1] * Kilometres7[i] + beta7[2] * Zones7[i] +
                    beta7[3] * Bonus7[i] + beta7[4] * Make7[i] + beta7[5] * Insured7[i] +
                    gamma7);
  y7[i] ~ poisson(lambda7);
}
  
  
}








generated quantities {
  // Posterior predictive checks for the main model
  int y_pred1[N1];
  int y_pred2[N2];
  int y_pred3[N3];
  int y_pred4[N4];
  int y_pred5[N5];
  int y_pred6[N6];
  int y_pred7[N7];



  for (i in 1:N1) {
    y_pred1[i] = poisson_rng(exp(alpha1 + beta1[1] * Kilometres1[i] + beta1[2] * Zones1[i] +
                    beta1[3] * Bonus1[i] + beta1[4] * Make1[i] + beta1[5] * Insured1[i] +
                                gamma1));
  }

  for (i in 1:N2) {
    y_pred2[i] = poisson_rng(exp(alpha2 + beta2[1] * Kilometres2[i] + beta2[2] * Zones2[i] +
                                beta2[3] * Bonus2[i] + beta2[4] * Make2[i] + beta2[5] * Insured2[i] +
                                gamma2));
  }
  
  for (i in 1:N3) {
    y_pred3[i] = poisson_rng(exp(alpha3 + beta3[1] * Kilometres3[i] + beta3[2] * Zones3[i] +
                                beta3[3] * Bonus3[i] + beta3[4] * Make3[i] + beta3[5] * Insured3[i] +
                                gamma3));
  }

  for (i in 1:N4) {
    y_pred4[i] = poisson_rng(exp(alpha4 + beta4[1] * Kilometres4[i] + beta4[2] * Zones4[i] +
                                beta4[3] * Bonus4[i] + beta4[4] * Make4[i] + beta4[5] * Insured4[i] +
                                gamma4));
  }

  for (i in 1:N5) {
    y_pred5[i] = poisson_rng(exp(alpha5 + beta5[1] * Kilometres5[i] + beta5[2] * Zones5[i] +
                                beta5[3] * Bonus5[i] + beta5[4] * Make5[i] + beta5[5] * Insured5[i] +
                                gamma5));
  }
  
  for (i in 1:N6) {
  y_pred6[i] = poisson_rng(exp(alpha6 + beta6[1] * Kilometres6[i] + beta6[2] * Zones6[i] +
                                beta6[3] * Bonus6[i] + beta6[4] * Make6[i] + beta6[5] * Insured6[i] +
                                gamma6));
  }
    
  for (i in 1:N7) {
  y_pred7[i] = poisson_rng(exp(alpha7 + beta7[1] * Kilometres7[i] + beta7[2] * Zones7[i] +
                                beta7[3] * Bonus7[i] + beta7[4] * Make7[i] + beta7[5] * Insured7[i] +
                                gamma7));
  }
  
}

'
  
  
  
#beta[1] = Kilometres
#beta[2] = Zone
#beta[3] = Bonus
#beta[4] = Make
#beta[5] = Insured

  
# Ensure Zones variables are numeric
Zone1$Zone <- as.numeric(Zone1$Zone)
Zone2$Zone <- as.numeric(Zone2$Zone)
Zone3$Zone <- as.numeric(Zone3$Zone)
Zone4$Zone <- as.numeric(Zone4$Zone)
Zone5$Zone <- as.numeric(Zone5$Zone)
Zone6$Zone <- as.numeric(Zone6$Zone)
Zone7$Zone <- as.numeric(Zone7$Zone)


  
  
Frequencies1 <- Zone1$Claims
Frequencies2 <- Zone2$Claims
Frequencies3 <- Zone3$Claims
Frequencies4 <- Zone4$Claims
Frequencies5 <- Zone5$Claims
Frequencies6 <- Zone6$Claims
Frequencies7 <- Zone7$Claims

Kilometres1 <- Zone1$Kilometres
Kilometres2 <- Zone2$Kilometres
Kilometres3 <- Zone3$Kilometres
Kilometres4 <- Zone4$Kilometres
Kilometres5 <- Zone5$Kilometres
Kilometres6 <- Zone6$Kilometres
Kilometres7 <- Zone7$Kilometres

Zones1 <- Zone1$Zone
Zones2 <- Zone2$Zone
Zones3 <- Zone3$Zone
Zones4 <- Zone4$Zone
Zones5 <- Zone5$Zone
Zones6 <- Zone6$Zone
Zones7 <- Zone7$Zone


Bonus1 <- Zone1$Bonus
Bonus2 <- Zone2$Bonus
Bonus3 <- Zone3$Bonus
Bonus4 <- Zone4$Bonus
Bonus5 <- Zone5$Bonus
Bonus6 <- Zone6$Bonus
Bonus7 <- Zone7$Bonus


Make1 <- Zone1$Make
Make2 <- Zone2$Make
Make3 <- Zone3$Make
Make4 <- Zone4$Make
Make5 <- Zone5$Make
Make6 <- Zone6$Make
Make7 <- Zone7$Make


Insured1 <- Zone1$Insured
Insured2 <- Zone2$Insured
Insured3 <- Zone3$Insured
Insured4 <- Zone4$Insured
Insured5 <- Zone5$Insured
Insured6 <- Zone6$Insured
Insured7 <- Zone7$Insured


# Compile the Stan model

data_list <- list(
  N1 = length(Frequencies1),
  N2 = length(Frequencies2),
  N3 = length(Frequencies3),
  N4 = length(Frequencies4),
  N5 = length(Frequencies5),
  N6 = length(Frequencies6),
  N7 = length(Frequencies7),
  y1 = Frequencies1,
  y2 = Frequencies2,
  y3 = Frequencies3,
  y4 = Frequencies4,
  y5 = Frequencies5,
  y6 = Frequencies6,
  y7 = Frequencies7,
  Kilometres1 = Kilometres1,
  Kilometres2 = Kilometres2,
  Kilometres3 = Kilometres3,
  Kilometres4 = Kilometres4,
  Kilometres5 = Kilometres5,
  Kilometres6 = Kilometres6,
  Kilometres7 = Kilometres7,
  Zones1 = Zones1,
  Zones2 = Zones2,
  Zones3 = Zones3,
  Zones4 = Zones4,
  Zones5 = Zones5,
  Zones6 = Zones6,
  Zones7 = Zones7,
  Bonus1 = Bonus1,
  Bonus2 = Bonus2,
  Bonus3 = Bonus3,
  Bonus4 = Bonus4,
  Bonus5 = Bonus5,
  Bonus6 = Bonus6,
  Bonus7 = Bonus7,
  Make1 = Make1,
  Make2 = Make2,
  Make3 = Make3,
  Make4 = Make4,
  Make5 = Make5,
  Make6 = Make6,
  Make7 = Make7,
  Insured1 = Insured1,
  Insured2 = Insured2,
  Insured3 = Insured3,
  Insured4 = Insured4,
  Insured5 = Insured5,
  Insured6 = Insured6,
  Insured7 = Insured7
)



# Define K (number of predictors)
K <- 5  # You need to set this to the actual number of predictors in your model

# Update the data_list with K
data_list$K <- K

# Compile the Stan model
stan_model_2 <- stan_model(model_code = stan_code_2)

# Run the sampling
stan_fit_2 <- sampling(stan_model_2, data = data_list, chains = 2, warmup = 10000, iter = 20000)
print(stan_fit_2)






```

# Try using different package brms...

```{r}
# Install and load the brms package
#install.packages("brms")
library(brms)



# Ensure Zones variables are numeric
Zone1$Zone <- as.numeric(Zone1$Zone)
Zone2$Zone <- as.numeric(Zone2$Zone)
Zone3$Zone <- as.numeric(Zone3$Zone)
Zone4$Zone <- as.numeric(Zone4$Zone)
Zone5$Zone <- as.numeric(Zone5$Zone)
Zone6$Zone <- as.numeric(Zone6$Zone)
Zone7$Zone <- as.numeric(Zone7$Zone)
  
  
Frequencies1 <- Zone1$Claims
Frequencies2 <- Zone2$Claims
Frequencies3 <- Zone3$Claims
Frequencies4 <- Zone4$Claims
Frequencies5 <- Zone5$Claims
Frequencies6 <- Zone6$Claims
Frequencies7 <- Zone7$Claims

Insured1 <- Zone1$Insured
Insured2 <- Zone2$Insured
Insured3 <- Zone3$Insured
Insured4 <- Zone4$Insured
Insured5 <- Zone5$Insured
Insured6 <- Zone6$Insured
Insured7 <- Zone7$Insured


# Compile the Stan model

data_list <- list(
  N1 = length(Frequencies1),
  N2 = length(Frequencies2),
  N3 = length(Frequencies3),
  N4 = length(Frequencies4),
  N5 = length(Frequencies5),
  N6 = length(Frequencies6),
  N7 = length(Frequencies7),
  y1 = Frequencies1,
  y2 = Frequencies2,
  y3 = Frequencies3,
  y4 = Frequencies4,
  y5 = Frequencies5,
  y6 = Frequencies6,
  y7 = Frequencies7,
  Insured1 = Insured1,
  Insured2 = Insured2,
  Insured3 = Insured3,
  Insured4 = Insured4,
  Insured5 = Insured5,
  Insured6 = Insured6,
  Insured7 = Insured7
)


all_data <- rbind(
  data.frame(Zone = Zone1$Zone, Insured = Zone1$Insured, Claims = Zone1$Claims),
  data.frame(Zone = Zone2$Zone, Insured = Zone2$Insured, Claims = Zone2$Claims),
  data.frame(Zone = Zone3$Zone, Insured = Zone3$Insured, Claims = Zone3$Claims),
  data.frame(Zone = Zone4$Zone, Insured = Zone4$Insured, Claims = Zone4$Claims),
  data.frame(Zone = Zone5$Zone, Insured = Zone5$Insured, Claims = Zone5$Claims),
  data.frame(Zone = Zone6$Zone, Insured = Zone6$Insured, Claims = Zone6$Claims),
  data.frame(Zone = Zone7$Zone, Insured = Zone7$Insured, Claims = Zone7$Claims)
)

# Ensure the 'Zone' variable is numeric
all_data$Zone <- as.numeric(all_data$Zone)

# Fit the Bayesian regression model
brm_fit <- brm(Claims ~ Insured + (1|Zone), data = all_data, family = poisson(), chains = 2, iter = 3000 )

# Print a summary of the model
summary(brm_fit)





```

```{r}

library(bayesplot)
library(brms)


# Trace plots
mcmc_trace(brm_fit, regex_pars = c("Intercept", "Insured", "^Zone"))

# Autocorrelation plots
mcmc_autocorr(brm_fit, regex_pars = c("Intercept", "Insured", "^Zone"))



```

# Calculate MSE:

```{r}


#calcualte Bayesian




#Calculate sample mean






```

# Posterior Predictive Check:

```{r}



library(bayesplot)

# Extract simulated data from the stan_fit object
yrep1 <- as.array(stan_fit, "y_rep1")
yrep2 <- as.array(stan_fit, "y_rep2")
yrep3 <- as.array(stan_fit, "y_rep3")
yrep4 <- as.array(stan_fit, "y_rep4")
yrep5 <- as.array(stan_fit, "y_rep5")
yrep6 <- as.array(stan_fit, "y_rep6")
yrep7 <- as.array(stan_fit, "y_rep7")

Frequencies1
Frequencies2
Frequencies3
Frequencies4
Frequencies5
Frequencies6
Frequencies7

yrep1fixed <- yrep1[1:6, 1,]
ppc_hist(Frequencies1, yrep1fixed, binwidth = 50)
ppc_dens_overlay(Frequencies1, yrep1fixed, binwidth = 50)

```

\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~

\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~\~

# Claim Severity Gamma Hierarchical Model:

```{r}

#claim severity follows a Gamma distribution which shape and scale parameters drawn from two other distributions

#Stan code used to fit and find parameters for the hierarchical model:

library(rstan)


stan_code_2 <- '

data {

  int<lower=0> N1; //# observations in Zone 1
  int<lower=0> N2; 
  int<lower=0> N3; 
  int<lower=0> N4; 
  int<lower=0> N5; 
  int<lower=0> N6; 
  int<lower=0> N7; 
  
  real<lower=0> y1[N1]; //payments in Zone 1
  real<lower=0> y2[N2]; 
  real<lower=0> y3[N3]; 
  real<lower=0> y4[N4]; 
  real<lower=0> y5[N5]; 
  real<lower=0> y6[N6]; 
  real<lower=0> y7[N7]; 
  
  
}



parameters {

  real<lower=0> theta; 
  real<lower=0> tau;

  real<lower=0> alpha1; //alpha parameter for Zone 1
  real<lower=0> alpha2; // 
  real<lower=0> alpha3; 
  real<lower=0> alpha4; 
  real<lower=0> alpha5; 
  real<lower=0> alpha6; 
  real<lower=0> alpha7; 
  
  real<lower=0> beta;  //shared rate for gamma distribution

  
}


model {


  //priors
  theta ~ uniform(.9, 1.1); // 
  tau ~ uniform(1,2);  // 
  
  
  beta ~ uniform(.0000005,.00005);
  

  //hierarchical model

  alpha1 ~ gamma(theta,tau);
  alpha2 ~ gamma(theta,tau);
  alpha3 ~ gamma(theta,tau);
  alpha4 ~ gamma(theta,tau);
  alpha5 ~ gamma(theta,tau);
  alpha6 ~ gamma(theta,tau);
  alpha7 ~ gamma(theta,tau);
  
  
  //likelihoods
  y1 ~ gamma(alpha1, beta);
  y2 ~ gamma(alpha2, beta);
  y3 ~ gamma(alpha3, beta);
  y4 ~ gamma(alpha4, beta);
  y5 ~ gamma(alpha5, beta);
  y6 ~ gamma(alpha6, beta);
  y7 ~ gamma(alpha7, beta);
  
}


'
  
  
  
Payment1 <- Zone1$Payment
Payment2 <- Zone2$Payment
Payment3 <- Zone3$Payment
Payment4 <- Zone4$Payment
Payment5 <- Zone5$Payment
Payment6 <- Zone6$Payment
Payment7 <- Zone7$Payment


  

stan_model_2 <- stan_model(model_code = stan_code_2)

data_list_2 <- list(

  N1 = length(Payment1),
  y1 = Payment1,
  N2 = length(Payment2),
  y2 = Payment2,
  N3 = length(Payment3),
  y3 = Payment3,
  N4 = length(Payment4),
  y4 = Payment4,
  N5 = length(Payment5),
  y5 = Payment5,
  N6 = length(Payment6),
  y6 = Payment6,
  N7 = length(Payment7),
  y7 = Payment7
)


# Set wider initial values for parameters for each chain
init_values <- list(
  list(theta = 1, tau = 1, alpha1 = 1, alpha2 = 1, alpha3 = 1, alpha4 = 1, alpha5 = 1, alpha6 = 1, alpha7 = 1, beta = 0.000001),
  list(theta = 1, tau = 1, alpha1 = 1, alpha2 = 1, alpha3 = 1, alpha4 = 1, alpha5 = 1, alpha6 = 1, alpha7 = 1, beta = 0.000001),
  list(theta = 1, tau = 1, alpha1 = 1, alpha2 = 1, alpha3 = 1, alpha4 = 1, alpha5 = 1, alpha6 = 1, alpha7 = 1, beta = 0.000001),
  list(theta = 1, tau = 1, alpha1 = 1, alpha2 = 1, alpha3 = 1, alpha4 = 1, alpha5 = 1, alpha6 = 1, alpha7 = 1, beta = 0.000001)
)

# Run the sampling with specified initial values
stan_fit_2 <- sampling(stan_model_2, data = data_list_2, chains = 4, iter = 20000, warmup = 10000, init = init_values)

# Print the results
print(stan_fit_2)



```

```{r}


library(rstan)

plot(stan_fit)
traceplot(stan_fit)
summary(stan_fit)


library(bayesplot)
autocorrelation_plot <- mcmc_acf(stan_fit)
print(autocorrelation_plot)


```

# Gamma Regression Model for Claim Severity:

Stan Code:

```{r}

library(rstan)


Zone3


stan_code <- "
data {
  int<lower=0> N;          // Number of observations
  int<lower=0> p;          // Number of predictors
  vector[N] Z;             // Claim severity data
  matrix[N, p] X;          // Design matrix
}

parameters {
  real alpha;              // Intercept
  vector[p] beta;          // Coefficients for predictors
  real<lower=0> nu;         // Shape parameter of the gamma distribution
  vector<lower=0>[p] kappa; // Shape parameter for the Inverse Gamma prior on nu
}

model {
  vector[N] mu;            // Mean parameter
  
  
  // Log-likelihood of Inverse Gamma prior for nu
  //nu ~ inv_gamma(0.001, 0.001);
  
  // Non-informative uniform priors on regression coefficients
  alpha ~ uniform(-10, 10); 
  beta ~ uniform(-10, 10);  
  
  // Prior on the shape parameter for each coefficient
  for (i in 1:p) {
  kappa[i] ~ inv_gamma(0.001, 0.001);
  }
  
  
  // Gamma likelihood
  Z ~ gamma(nu, exp(alpha + dot_product(X, beta)) / nu);
  
  
  
}
  

generated quantities {
  vector[N] y_pred;
  
  for (i in 1:N)
    y_pred[i] = gamma_rng(nu, exp(alpha + X[i] * beta) / nu);
}
"
  


#Data
set.seed(123)
data <- list(N = 315,
             p = 4,
             Z = Zone3$Payment,
             X = cbind(Zone3$Kilometres, Zone3$Bonus, Zone3$Make, Zone3$Insured))


# Compile the Stan model
stan_model <- stan_model(model_code = stan_code)


## Specify different initial values
init_values <- list(
  list(alpha = 0.1, beta = rep(0.1, data$p), nu = 2, kappa = rep(0.1, data$p)),
  list(alpha = -0.5, beta = rep(-0.5, data$p), nu = 1, kappa = rep(0.1, data$p)),
  
)


# Fit the model using MCMC
fit <- sampling(stan_model, data = data, chains = 4, iter = 2000, warmup = 1000)

# Display summary of the fitted model
print(fit)




```

# Claim Severity OLS Frequentist Model:

# Test for claim frequency with brms package:

```{r}

library(bpr)
library(lme4)
library(coda)
library(brms)
library(ggplot2)
library(bayesplot)


# Ensure Zones variable is numeric
Swedish$Zone <- as.numeric(Swedish$Zone)

# Fit the Poisson hierarchical regression model with brms
model <- brm(Claims ~ Insured + (1 | Zone), data = Swedish, family = poisson(), chains = 3, iter = 1000)


# Extract posterior samples
samples <- posterior_samples(model)

# Trace plot
mcmc_trace(samples)

# Autocorrelation plot
mcmc_acf(samples)



```
